Eye in a Disk: eyeIntegration human pan-eye and body transcriptome database version 1.00

### 2019-02-14 manuscript.Rmd

## Differential Gene Expression Analysis and GO term enrichment
We used the cleaned data to determine differential gene expression between different sub-tissue types. First, we generated a synthetic body set to serve as single representative sub-tissue type for pan-body gene expression by randomly sampling GTEx tissues.  We used the voom function from the Limma R package to convert gene expression to precision weights, and then performed pairwise differential expression tests for all combinations of sub-tissues using an empirical Bayes test [@ritchieLimmaPowersDifferential2015; @lawVoomPrecisionWeights2014]. We extracted significant genes (FDR p < 0.01) for all `r limma_DE_tests %>% nrow()` comparisons and used these to calculate GO enrichment. The significant gene list for each eye sub-tissue was split into upregulated and down regulated sets and each set was tested for enrichment using the enrichGO function from the clusterProfiler R package (q-value < 0.01) [@yuClusterProfilerPackageComparing2012].

@@

## Differential expression and gene ontology enrichment tests allow quick comparison of gene differences between groups
Using our expression data, we performed multiple differential comparisons at the sub-tissue level within all eye tissues and against a pan-body synthetic set comprised of a stratified sample of all tissues present in our subset of the GTEx dataset, allowing users to quickly identify eye specific genes across `r limma_DE_tests %>% nrow()` different comparisons.  The user can view the results under the ‘Differential’ selection under the ‘expression’ tab (`r paste0(fig_cap(name = 'appScreenshots', display = 'cite'),'F')`).  As with expression, the user can select which version of the web app to draw data from as well as select for gene or transcript level comparisons. The user additionally has the option to select different gene classes to examine, protein coding, lincRNA etc.

>>>>

## Differential Gene Expression Analysis and GO term enrichment
We used the cleaned data to determine differential gene expression between different sub-tissue types. First, we generated a synthetic body set to serve as single representative sub-tissue type for pan-body gene expression by randomly sampling GTEx tissues.  We used the voom function from the Limma R package to convert gene expression to precision weights, and then performed pairwise differential expression tests for all combinations of eye sub-tissues, the synthetic body tissue, and human body tissues using an empirical Bayes test [@ritchieLimmaPowersDifferential2015; @lawVoomPrecisionWeights2014]. We extracted significant genes (FDR p < 0.01) for all `r limma_DE_tests %>% nrow()` comparisons and used these to calculate GO enrichment. The significant gene list for each eye sub-tissue was split into upregulated and down regulated sets and each set was tested for enrichment using the enrichGO function from the clusterProfiler R package (q-value < 0.01) [@yuClusterProfilerPackageComparing2012].

@@

## Differential expression and gene ontology enrichment tests allow quick comparison of gene differences between groups
Using our expression data, we performed multiple differential comparisons at the sub-tissue level within all eye tissues and against a pan-body synthetic set comprised of a stratified sample of all tissues present in our subset of the GTEx dataset, allowing users to quickly identify eye specific genes across `r limma_DE_tests %>% nrow()` different comparisons. We have expanded the differential tests in the 2019 EiaD by adding the GTEx tissues as direct comparisons to our eye sub-tissues. The user can view the results under the ‘Differential’ selection under the ‘expression’ tab (`r paste0(fig_cap(name = 'appScreenshots', display = 'cite'),'F')`).  As with expression, the user can select which version of the web app to draw data from as well as select for gene or transcript level comparisons. The user additionally has the option to select different gene classes to examine, protein coding, lincRNA etc.
